Exploring probability distributions for bivariate temporal granularities

### 2019-11-29 main.Rmd
cyclic intro change

Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra[@Ning2002-tf]. It is assumed that there exists a “bottom” granularity and Calendar algebra operations are designed to generate
new granularities from the bottom one or recursively, from those already generated.
Thus, the relationship between the operand(s) and the resulting granularities are encoded
in the operations.

@@

@@

@@

Periodicity is very common in all kinds of data. Time granularities which can accomodate periodicities can be constructed by relating two linear granularities. The mappings between linear granularities can be regular or irregular. For example, a regular mapping exists between minutes and hours, where 60 minutes always add up to 1 hour. On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31. Hence, time granularities can also be **circular** or **aperiodic** depending on if the mapping of the linear granularities is regular or irregular. Examples of circular time granularities include hour of the day, and day of the week, whereas examples for aperiodic time granularities can be day of the month or public holidays.

In a cyclic organization of time, the domain is composed of a set of recurring time values. Hence, any time value can be proceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday). Intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.

We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data. A time domain, as defined by Bettini, is essentially a mapping of the index set to the time index of a tsibble. A linear granularity is a mapping of the index set to subsets of the time domain. For example, if the time index is days, then a linear granularity might be weeks, months or years. Lining up with the the assumption in [@Bettini2000-vy] that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists and is represented by the index of the tsibble.

The mappings of two linear granularities forming a cyclic granularity can be regular or irregular. It will be regular if, for example, they are formed by grouping operation. Grouping operation would ensure that each granule of the resulting granularity consists of same number of granules of bottom granularity. However, if they are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of bottom granularity. The relationship can still be periodic, depending on the choice of periods and hence these are also useful in exploring periodicities. However, the formalism would differ for circular (regular mapping) and aperiodic (irregular mapping) granularities.

>>>>

Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf]. It is assumed that there exists a “bottom” granularity and Calendar algebra operations are designed to generate
new granularities from the bottom one or recursively, from those already generated.
Thus, the relationship between the operand(s) and the resulting granularities are encoded
in the operations.

@@

**Examples** :

@@

For more variations of grouping operations and granule oriented operations, the readers are remcommended to see [@Ning2002-tf].s

@@

<!-- sub: regular and irregular mapping leads to circular and quasi-circular granularities -->

Periodicity is very common in all kinds of data. Time granularities which can accomodate periodicities can be constructed by relating two linear granularities. The mappings between linear granularities can be regular or irregular. For example, a regular mapping exists between minutes and hours, where 60 minutes always add up to 1 hour. On contrary, an irregular mapping exists between days and months, since one might can have days ranging from 28 to 31. Hence, time granularities can also be **circular** or **aperiodic** depending on if the mapping of the linear granularities is regular or irregular. Mappings will be regular if, for example, they are formed by grouping operation. Grouping operation would ensure that each granule of the resulting granularity consists of same number of granules of bottom granularity. However, if they are formed by the altering-tick operation, for example, granules of the resulting granularity is composed of different number of granules of bottom granularity. However, the formalism would differ for circular (regular mapping) and quasi-circular (irregular mapping) granularities. Examples of circular granularities include hour of the day, and day of the week, whereas examples for quasi-circular granularities can be day of the month or public holidays.

<!--sub:How to relate linear granularities to tsibble and manipulation of indices required for circular and quasi-circular granularities -->

We propose a formalism of cyclic time granularities through the tsibble [@wang2019tsibble] framework of organizing temporal data. A time domain, as defined by Bettini, is essentially a mapping of the index set to the time index of a tsibble. A linear granularity is a mapping of the index set to subsets of the time domain. For example, if the time index is days, then a linear granularity might be weeks, months or years. Lining up with the the assumption in [@Bettini2000-vy] that all linear granularities can be generated from the bottom granularity by calendar algebraic operations, we assume that a bottom granularity exists and is represented by the index of the tsibble. In a cyclic organization of time, the domain is composed of a set of recurring time values, where any time value can be proceeded and succeeded by another time value (for e.g Monday comes before Wednesday but Monday also succeeds Wednesday). Hence intuitively, cyclic granularities are additional abstractions of time which are not linear and hence the index order of linear granularities needs to be manipulated so that it leads to repetitive categorization of time.
